---
title: "qPCR_outcomes"
author: "LRA"
date: "5/4/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
region_order <- c("AK","BC","WA","OR","BB","SD")
det_order <- c("Negative","Positive","Low")

```

### qPCR outcomes from 2019 NSF samples

NSF lesion and green tissue samples were collected in July/Aug of 2019 and stored in 70% ethanol.

I extracted and analyzed for qPCR 60 lesion samples from 4 regions (AK, WA, OR, BB).

Positive results (e.g. Lz DNA in the samples) were low: 2/15 in AK, 10/15 in WA, 2/20 in OR, 1/10 in BB.

Good news is that the range of cell counts in other regions fell within the range found in WA.

Note, the "Low" category is for positive results with <1 cell / mg.

```{r data,echo=FALSE,warning=FALSE}
qPCR <- read.csv("qPCR_detection_04_2020.csv")

run <- subset(qPCR,Run=="Y")
run$Region <- ordered(run$Region,levels=region_order)
run$Detection <- ordered(run$Detection,levels=det_order)
counts <- run %>%
  group_by(Region,Detection)%>%
  summarise(total=length(Run))

ggplot(run,aes(x=Detection))+geom_histogram(aes(fill=SiteCode),stat = "count",color="black")+
  facet_wrap(~Region)+
  theme_bw()+
  xlab("qPCR detection results")+
  ylab("Number of tissue samples")+
  labs(title="qPCR results from NSF tissue")+
  scale_y_continuous(limits = c(0,20))+
  geom_text(data=counts,aes(x=Detection,y=19.5,label=total))

ggplot(data=run,aes(x=Region,y=Cells,color=Detection))+geom_point()+
  theme_bw()+
  xlab("Region")+
  ylab("Lz cells / mg wet tissue")+
  labs(title="Lz cell counts from positive lesion tissue samples")
```

### Possible reasons for low rate of positives:

(1) Sample selection - tissue samples were likely not optimal as partners have little experience in selecting lesions. 

(2) Ethanol storage - it's possible that the ethanol storage did not preserve DNA well, leading to low yields in the extracts. Morgan and I ran a methods test in September, by pairing samples from the same lesion and flash freezing half while storing the other half in ethanol. There was no clear difference, but we again had a low rate of positives with the tests. 

(3) Primer specificity? We've mentioned this a few times, my understanding is that there might be adjustments to either the primers or to the qPCR method (e.g. using SYBR Green?) that might lead to more success.

### Moving forward - 2019

For purposes of the NSF analysis, we want to be able to say we have Lz in all regions and ideally in all meadows where we find lesions. We can probably get away with saying that we confirmed Lz in all regions, but the figures above are not very convincing (especially when we have 10:1 ratio of negatives to positives).

One strategy is that we can run more samples. We collected 12 tissue samples per meadow (site). For the first round of qPCR anaylsis, I extracted and analyzed a subset of samples (no more than 4 per meadow). So, theoretically, we have more tissue samples to run at each site (up to 8 more per site). However, I tried to select the most promising lesion tissue samples, based on the scanned images. So the remaining tissues samples may be slightly worse candidates for qPCR than those already run. We'd also have to think about the expense of running many more samples in pursuit of a handful of positives. 

Regarding (2) above, we planned to run tests in Colleen's lab to confirm the presence of amplifiable DNA in the extracts. Basically, we want to make sure that the ethanol storage isn't compromising DNA extraction. The plan was to run universal eukaryotic primers to confirm there is any DNA in the samples (eelgrass or Lz). This would eliminate sample collection and extraction as sources of error.

Assuming we do have DNA in the extracts, we could do further methods development (e.g. address (3) above). This would likely need to happen in Colleen's lab, with Colleen or Tori assisting Lillian :)

Note, I am currently planning to extract and run additional 2019 samples from the remaining regions (BC, SD, more BB) in June 2020. I expect to have low success rates with these samples as well. Potentially, we could think about adjusting the extraction methods - possibly using a different kit? When we last talked it didn't sound like this would be very fruitful, but we could try it. Not sure if there are any other changes we could make, with the equipment at FHL, to improve success rates.

### Moving forward - 2020

We need to decide on our sampling plan for summer 2020 within the next few weeks. A few possibilities:

(1) Ask partners to take the same approach as last year - e.g. identify lesions in the field, store in ethanol.

(2) Ask partners to ship us fresh eelgrass from the field so we can identify lesions and flash freeze, rather than use ethanol.

(3) Possibly use an alternative preservation, such as drying in silica?

One issue with choosing a new method is that we don't actually know what/if there's an issue with the 2019 method. We likely won't have any additional information to make that decision. Possibly sometime this summer we will be able to do the tests to confirm that the ethanol preservation is okay, but we can't do those tests under current conditions.

Given this situation, my suggestion is that we ask partners to collect lesion samples at a subset of sites using the same methods as last year. At least it will be comparable, but we can reduce the effort if it turns out that there is an issue with the preservation.

A final thought - Deanna (Jay's postdoc) has mentioned that she has tissue leftover from the microbiome sampling (those tissue samples were preserved in Zymo buffer, and she analyzed the solution, not the tissue). We could investigate if the Zymo buffer is compatible with the kit extraction, and perhaps try running some 2019 microbiome tissue samples for qPCR.