---
title: "qPCR_outcomes_v2"
author: "LRA"
date: "8/27/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
region_order <- c("AK","BC","WA","OR","BB","SD")
det_order <- c("Negative","Positive","Low")
```

## Updated qPCR outcomes from 2019 and 2020 samples

### 2019 NSF Samples

NSF lesion and green tissue samples were collected in July/Aug of 2019 and stored in 70% ethanol.

I extracted and analyzed for qPCR between 24-33 lesion samples for each region.

Some samples had very low Lz cell counts (<1 cell/mg tissue). I excluded these low values from the positives as not biologically meaningful.

Lesion tissue tested positive for Lz cells in 51% of WA samples but positives were 0-15% of samples at other sites.

```{r lesion_2019, echo=FALSE,warning=FALSE}
qPCR <- read.csv("qPCR_detection_08_2020.csv")

run <- subset(qPCR,Run=="Y")
run$Region <- ordered(run$Region,levels=region_order)
run$Detection <- ordered(run$Detection,levels=det_order)
run_19 <-subset(run,Year=="2019")
run_20 <- subset(run,Year=="2020")

run_L <- subset(run_19,Tissue=="lesion")
counts_L <- run_L %>%
  group_by(Region,Detection)%>%
  summarise(total=length(Run))
plot_L <- ggplot(run_L,aes(x=Detection))+geom_histogram(aes(fill=SiteCode),stat = "count",color="black")+
  facet_wrap(~Region)+
  theme_bw()+
  xlab("qPCR detection results")+
  ylab("Number of tissue samples")+
  labs(title="qPCR results from NSF lesion tissue - 2019")+
  scale_y_continuous(limits = c(0,33))+
  scale_fill_brewer(palette="Oranges")+
  geom_text(data=counts_L,aes(x=Detection,y=32,label=total))

plot_L
```

After running the lesion tissue, I identified the paired green tissue samples for any lesions that were positive for Lz, plus 10 additional green tissue samples from AK. I extracted and ran a total of 35 green tissue samples. 

A fraction of the green tissue samples did show a positive presence of Lz, but at very low quantities (<1 cell/mg tissue). I again excluded these low results as not biologically meaningful, leaving only 2 true positives from the green tissue.

```{r green_2019, echo=FALSE,warning=FALSE}
run_G <- subset(run_19,Tissue=="green")
counts_G <- run_G %>%
  group_by(Region,Detection)%>%
  summarise(total=length(Run))

plot_G <- ggplot(run_G,aes(x=Detection))+geom_histogram(aes(fill=SiteCode),stat = "count",color="black")+
  facet_wrap(~Region)+
  theme_bw()+
  xlab("qPCR detection results")+
  ylab("Number of tissue samples")+
  labs(title="qPCR results from NSF green tissue - 2019")+
  scale_fill_brewer(palette="Greens")+
  scale_y_continuous(limits = c(0,17.5))+
  geom_text(data=counts_G,aes(x=Detection,y=17,label=total))

plot_G
```

Note, no green tissue was processed for SD because no lesion tissue tested positive for Lz for SD.

#### Take-aways from the 2019 samples:

(1) Lesion selection is a major factor in finding lesions that test positive for Lz. Even in WA, with more experience in lesion selection, only ~50% of samples were positive.

(2) There may be geographic effects. No positives were found in any site in the SD region, and in other regions besides WA, positives were found at only 1 or 2 sites. 

(3) Key question - does the qPCR test result represent an _active_ infection? I.e. lesion tissue that tests negative may still have had Lz present to stimulate the visible plant immune response, but the Lz are no longer present in the tissue. This would mean the qPCR assay can detect active infection but can't rule out previously active infection.

### 2020 NSF Samples

In 2020, we asked Bodega and San Diego to ship fresh tissue to FHL so we could optimize lesion selection and preservation. We targeted lesions from Leaf 2 (youngest and freshest tissue) and we targeted lesions with a clear dark edge and evidence of rapid expansion (jagged lines and dark center).

I extracted and ran 15 lesion samples from Bodega sites (5 from each of 3 sites) and 5 lesion samples from San Diego (2 sites).

At BB-A, all 5 lesion tissue samples were negative. At BB-C, 2 tissue samples were positive. At BB-E, 3 tissue samples were positive. All 5 lesion tissue samples were negative at the San Diego sites.

```{r lesion_2020, echo=FALSE,warning=FALSE}
run_20_L <- subset(run_20,Tissue=="lesion")
counts_20L <- run_20_L %>%
  group_by(Region,Detection)%>%
  summarise(total=length(Run))
run_20_G <- subset(run_20,Tissue=="green")
counts_20G <- run_20_G %>%
  group_by(Region,Detection)%>%
  summarise(total=length(Run))

plot_20_L <- ggplot(run_20_L,aes(x=Detection))+geom_histogram(aes(fill=SiteCode),stat = "count",color="black")+
  facet_wrap(~Region)+
  theme_bw()+
  xlab("qPCR detection results")+
  ylab("Number of tissue samples")+
  labs(title="qPCR results from NSF tissue - 2020")+
  scale_y_continuous(limits = c(0,15))+
  scale_fill_brewer(palette = "Oranges")+
  geom_text(data=counts_20L,aes(x=Detection,y=14,label=total))
plot_20_L
plot_20_G <- ggplot(run_20_G,aes(x=Detection))+geom_histogram(aes(fill=SiteCode),stat = "count",color="black")+
  facet_wrap(~Region)+
  theme_bw()+
  xlab("qPCR detection results")+
  ylab("Number of tissue samples")+
  labs(title="qPCR results from NSF tissue - 2020")+
  scale_y_continuous(limits = c(0,10))+
  scale_fill_brewer(palette = "Greens")+
  geom_text(data=counts_20G,aes(x=Detection,y=9,label=total))
plot_20_G
```

#### Take-aways from the 2020 samples:

(1) With improved lesion selection, we can identify more positives, i.e. it was worth it to have Bodega ship us tissue.

(2) Even with optimal lesion selection, not all sites will show positives (e.g. BB-A). It's possible we could've gotten one or two positives from running additional BB-A samples, but that would still give us <20% positivity. Again, do we think Lz is not present at these sites? Or is there something about the dynamics of infection, e.g. Lz may leave the tissue rapidly after immune response? Plants in Bodega are also long-lived, so even Leaf 2 tissue may be weeks old. 

(3) Tissue from San Diego is questionable (Phyllospadix?) but we found no positives in either 2019 (n=26) or 2020 (n=5) samples.

### Repeated sampling at Fourth of July beach, WA

We did repeated sampling at Fourth of July to test the ethanol storage method and to improve the lab processing method. In Sept 2019, May 2020, and June 2020, we selected and preserved 15 lesions from Fourth of July. I extracted and ran these samples within days of sample preservation for the Sept 2019 and June 2020 samples and about 1 month after preservation for the May 2020 samples.

Note, paired green tissue from these samples is in storage, and can be processed if need be. 

```{r WA_tests,echo=FALSE,warning=FALSE}
tests <- read.csv("qPCR_WA_tests.csv")
tests$YearMonth <- paste(tests$Month,tests$Year,sep=" ")
tests$YearMonth <- ordered(tests$YearMonth,levels=c("September 2019","May 2020","June 2020"))
tests$Detection <- ordered(tests$Detection,levels=det_order)

count_T <- tests %>%
  group_by(YearMonth,Detection)%>%
  summarise(total=length(Run))

plot_T <- ggplot(tests,aes(x=Detection,fill=Detection))+geom_histogram(stat = "count")+
  facet_wrap(~YearMonth)+
  theme_bw()+
  xlab("qPCR detection results")+
  ylab("Number of tissue samples")+
  labs(title="qPCR results from Fourth of July")+
  scale_y_continuous(limits = c(0,16))+
  scale_color_brewer(palette='Dark2')+
  geom_text(data=count_T,aes(x=Detection,y=16,label=total))
plot_T

tests_P <- subset(tests,Detection=="Positive")

cells_T <- ggplot(data=tests_P,aes(x=YearMonth,y=Cells.mg,color=YearMonth))+geom_jitter()+
  stat_summary(fun.y=mean,geom="point",shape=8,size=2,position=position_dodge(width=0.75),
               show.legend = FALSE)+
  theme_bw()+
  #scale_color_brewer(palette='Dark2')+
  xlab("Sampling Date")+
  ylab("Lz cells / mg wet tissue")+
  labs(title = "Lz cell counts from positive lesions at Fourth of July",
       subtitle = "Star indicates mean cell count")+
  theme(legend.title = element_blank())
cells_T
```

#### Take-aways from WA tests

(1) Ethanol storage does not compromise cell counts

(2) Timing of sampling is important - does infection come in waves? High cell counts found in May 2020 were not repeated in June 2020, and overall infection severity was anecdotally lower (can verify this with EELISA later). Catching high cell counts and/or a high proportion of positives may depend on when sampling occurs relative to unknown infection dynamics.
