---
title: "SampleSizeAnalysisDiseaseBlades"
author: "LRA"
date: "4/22/2020"
output: github_document
fig_width: 10 
fig_height: 12
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
```

## Sample size for disease surveys

Can we reduce the number of blades collected for analysis of prevalence and severity?

First look at paired t-tests of Prevalence, Severity, and Blade Area. Compare transect-level means at each site, for the full sample size (20 blades) and a subsample (15, 10, or 5).

Note the n=5 blades are the epiphyte blades, so they are evenly distributed along the transects.

```{r data, echo=FALSE}
disease <- read.csv("disease_2019.csv")
t_20 <- disease %>%
  group_by(Region,SiteCode,TidalHeight,Transect)%>%
  summarise(Prevalence=mean(Prevalence),BladeArea=mean(BladeArea),
            Severity=mean(Severity),Count=length(Blade))
disease_15 <- subset(disease,Blade<16)
t_15 <- disease_15 %>%
  group_by(Region,SiteCode,TidalHeight,Transect)%>%
  summarise(Prevalence=mean(Prevalence),BladeArea=mean(BladeArea),
            Severity=mean(Severity),Count=length(Blade))
disease_10 <- subset(disease,Blade<11)
t_10 <- disease_10 %>%
  group_by(Region,SiteCode,TidalHeight,Transect)%>%
  summarise(Prevalence=mean(Prevalence),BladeArea=mean(BladeArea),
            Severity=mean(Severity),Count=length(Blade))
disease_5s <- subset(disease,Blade>15)
t_5s <- disease_5s %>%
  group_by(Region,SiteCode,TidalHeight,Transect)%>%
  summarise(Prevalence=mean(Prevalence),BladeArea=mean(BladeArea),
            Severity=mean(Severity),Count=length(Blade))

t_20 <- unite(t_20,col="TransectId",c("Region","SiteCode","TidalHeight","Transect"),sep="_",remove = FALSE)
t_15 <- unite(t_15,col="TransectId",c("Region","SiteCode","TidalHeight","Transect"),sep="_",remove = FALSE)
t_10 <- unite(t_10,col="TransectId",c("Region","SiteCode","TidalHeight","Transect"),sep="_",remove = FALSE)
t_5s <- unite(t_5s,col="TransectId",c("Region","SiteCode","TidalHeight","Transect"),sep="_",remove = FALSE)
t_20$TransectId <- as.factor(t_20$TransectId)
t_15$TransectId <- as.factor(t_15$TransectId)
t_10$TransectId <- as.factor(t_10$TransectId)
t_5s$TransectId <- as.factor(t_5s$TransectId)

transects <- levels(t_20$TransectId)

t_20 <- unite(t_20,col="MeadowId",c("Region","SiteCode"),sep="_",remove = FALSE)
t_15 <- unite(t_15,col="MeadowId",c("Region","SiteCode"),sep="_",remove = FALSE)
t_10 <- unite(t_10,col="MeadowId",c("Region","SiteCode"),sep="_",remove = FALSE)
t_5s <- unite(t_5s,col="MeadowId",c("Region","SiteCode"),sep="_",remove = FALSE)
t_20$MeadowId <- as.factor(t_20$MeadowId)
t_15$MeadowId <- as.factor(t_15$MeadowId)
t_10$MeadowId <- as.factor(t_10$MeadowId)
t_5s$MeadowId <- as.factor(t_5s$MeadowId)

meadows <- levels(t_20$MeadowId)

for(i in seq_along(meadows)){
  temp1 <- subset(t_20,MeadowId==meadows[i])
  temp2 <- subset(t_15,MeadowId==meadows[i])
  # if (is.na(mean(temp4$Ave_3reps)==TRUE)) next
  # else
  test <- t.test(temp1$Prevalence,temp2$Prevalence,paired=TRUE)
  if(i == 1){sig_table_15 <- data.frame(MeadowId=meadows)}else
  sig_table_15[i,"P-value"] <- test$p.value
}

for(i in seq_along(meadows)){
  temp1 <- subset(t_20,MeadowId==meadows[i])
  temp2 <- subset(t_15,MeadowId==meadows[i])
  testP <- t.test(temp1$Prevalence,temp2$Prevalence,paired=TRUE)
  testS <- t.test(temp1$Severity,temp2$Severity,paired=TRUE)
  testBA <- t.test(temp1$BladeArea,temp2$BladeArea,paired=TRUE)
  if(i == 1){sig_table_15 <- data.frame(MeadowId=meadows)}else
    sig_table_15[i,"Prevalence"] <- testP$p.value
  sig_table_15[i,"Severity"] <- testS$p.value
  sig_table_15[i,"Blade Area"] <- testBA$p.value
}

for(i in seq_along(meadows)){
  temp1 <- subset(t_20,MeadowId==meadows[i])
  temp2 <- subset(t_10,MeadowId==meadows[i])
  testP <- t.test(temp1$Prevalence,temp2$Prevalence,paired=TRUE)
  testS <- t.test(temp1$Severity,temp2$Severity,paired=TRUE)
  testBA <- t.test(temp1$BladeArea,temp2$BladeArea,paired=TRUE)
  if(i == 1){sig_table_10 <- data.frame(MeadowId=meadows)}else
    sig_table_10[i,"Prevalence"] <- testP$p.value
  sig_table_10[i,"Severity"] <- testS$p.value
  sig_table_10[i,"Blade Area"] <- testBA$p.value
}

for(i in seq_along(meadows)){
  temp1 <- subset(t_20,MeadowId==meadows[i])
  temp2 <- subset(t_5s,MeadowId==meadows[i])
  testP <- t.test(temp1$Prevalence,temp2$Prevalence,paired=TRUE)
  testS <- t.test(temp1$Severity,temp2$Severity,paired=TRUE)
  testBA <- t.test(temp1$BladeArea,temp2$BladeArea,paired=TRUE)
  if(i == 1){sig_table_5s <- data.frame(MeadowId=meadows)}else
    sig_table_5s[i,"Prevalence"] <- testP$p.value
  sig_table_5s[i,"Severity"] <- testS$p.value
  sig_table_5s[i,"Blade Area"] <- testBA$p.value
}
```

### P-values of paired t-tests for n=15
```{r fifteen,echo=FALSE}
print(sig_table_15)
```

### P-values of paired t-tests for n=10
```{r ten,echo=FALSE}
print(sig_table_10)
```

### P-values of paired t-tests for n=5
```{r five,echo=FALSE}
print(sig_table_5s)
```

Few significant tests, so statistically we are fairly safe to reduce sample size.

Now compare correlation between subsamples and full sample

```{r plots, echo=FALSE}
t_20$Replicates <- "Twenty"
t_15$Replicates <- "Fifteen"
t_10$Replicates <- "Ten"
t_5s$Replicates <- "Five"

full <- rbind(t_20,t_15,t_10,t_5s)

wide_P <- full[,c("TransectId","Replicates","Prevalence")]
wide_P <- spread(wide_P,key=Replicates,Prevalence)
wide_P <- separate(wide_P,col="TransectId",into=c("Region","SiteCode","TidalHeight","Transect"),sep="_",remove = FALSE)

wide_S <- full[,c("TransectId","Replicates","Severity")]
wide_S <- spread(wide_S,key=Replicates,Severity)
wide_S <- separate(wide_S,col="TransectId",into=c("Region","SiteCode","TidalHeight","Transect"),sep="_",remove = FALSE)

wide_BA <- full[,c("TransectId","Replicates","BladeArea")]
wide_BA <- spread(wide_BA,key=Replicates,BladeArea)
wide_BA <- separate(wide_BA,col="TransectId",into=c("Region","SiteCode","TidalHeight","Transect"),sep="_",remove = FALSE)

P15 <- ggplot(wide_P,aes(x=Twenty))+
  geom_point(aes(y=Fifteen,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  theme_bw()+
  labs(title="15 blades")+
  xlab("")+
  ylab("Prevalence")+
  theme(legend.position = "")

P10 <- ggplot(wide_P,aes(x=Twenty))+
  geom_point(aes(y=Ten,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  theme_bw()+
  labs(title="10 blades")+
  xlab("")+
  ylab("Prevalence")

P5 <- ggplot(wide_P,aes(x=Twenty))+
  geom_point(aes(y=Five,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
#  stat_smooth(method="lm",geom="smooth",aes(y=Five))+
  theme_bw()+
  labs(title="5 blades")+
  xlab("Prevalence of n=20 blades")+
  ylab("Prevalence")+
  theme(legend.position = "")

BA15 <- ggplot(wide_BA,aes(x=Twenty))+
  geom_point(aes(y=Fifteen,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  #stat_smooth(method="lm",geom="smooth",aes(y=Fifteen))+
  theme_bw()+
  xlab("")+
  labs(title="15 blades")+
  ylab("Blade Area")+
  theme(legend.position = "")

BA10 <- ggplot(wide_BA,aes(x=Twenty))+
  geom_point(aes(y=Ten,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  #stat_smooth(method="lm",geom="smooth",aes(y=Fifteen))+
  theme_bw()+
  labs(title="10 blades")+
  xlab("")+
  ylab("Blade Area")

BA5 <- ggplot(wide_BA,aes(x=Twenty))+
  geom_point(aes(y=Five,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  #stat_smooth(method="lm",geom="smooth",aes(y=Fifteen))+
  theme_bw()+
  xlab("Blade Area of n=20 blades")+
  ylab("Blade Area")+
  labs(title="5 blades")+
  theme(legend.position = "")


S15 <- ggplot(wide_S,aes(x=Twenty))+
  geom_point(aes(y=Fifteen,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  #stat_smooth(method="lm",geom="smooth",aes(y=Fifteen))+
  theme_bw()+
  xlab("")+
  ylab("Severity")+
  theme(legend.position = "")+
  labs(title="15 blades")

S10 <- ggplot(wide_S,aes(x=Twenty))+
  geom_point(aes(y=Ten,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  #stat_smooth(method="lm",geom="smooth",aes(y=Fifteen))+
  theme_bw()+
  xlab("")+
  ylab("Severity")+
  labs(title="10 blades")

S5 <- ggplot(wide_S,aes(x=Twenty))+
  geom_point(aes(y=Five,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  #stat_smooth(method="lm",geom="smooth",aes(y=Fifteen))+
  theme_bw()+
  xlab("Severity of n=20 blades")+
  ylab("Severity")+
  labs(title="5 blades")+
  theme(legend.position = "")
```

### Transect level plots

Prevalence

```{r P,echo=FALSE}
P15/P10/P5
```

Severity

```{r S,echo=FALSE}
S15/S10/S5
```

Blade Area

```{r BA,echo=FALSE}
BA15/BA10/BA5
```

```{r meadow,echo=FALSE}
m_20 <- disease %>%
  group_by(Region,SiteCode)%>%
  summarise(Prevalence=mean(Prevalence),BladeArea=mean(BladeArea),
            Severity=mean(Severity),Count=length(Blade))

m_15 <- disease_15 %>%
  group_by(Region,SiteCode)%>%
  summarise(Prevalence=mean(Prevalence),BladeArea=mean(BladeArea),
            Severity=mean(Severity),Count=length(Blade))

m_10 <- disease_10 %>%
  group_by(Region,SiteCode)%>%
  summarise(Prevalence=mean(Prevalence),BladeArea=mean(BladeArea),
            Severity=mean(Severity),Count=length(Blade))

m_5s <- disease_5s %>%
  group_by(Region,SiteCode)%>%
  summarise(Prevalence=mean(Prevalence),BladeArea=mean(BladeArea),
            Severity=mean(Severity),Count=length(Blade))

m_20 <- unite(m_20,col="MeadowId",c("Region","SiteCode"),sep="_",remove = FALSE)
m_15 <- unite(m_15,col="MeadowId",c("Region","SiteCode"),sep="_",remove = FALSE)
m_10 <- unite(m_10,col="MeadowId",c("Region","SiteCode"),sep="_",remove = FALSE)
m_5s <- unite(m_5s,col="MeadowId",c("Region","SiteCode"),sep="_",remove = FALSE)
m_20$MeadowId <- as.factor(m_20$MeadowId)
m_15$MeadowId <- as.factor(m_15$MeadowId)
m_10$MeadowId <- as.factor(m_10$MeadowId)
m_5s$MeadowId <- as.factor(m_5s$MeadowId)
m_20$Replicates <- "Twenty"
m_15$Replicates <- "Fifteen"
m_10$Replicates <- "Ten"
m_5s$Replicates <- "Five"

full_m <- rbind(m_20,m_15,m_10,m_5s)

wide_m_P <- full_m[,c("MeadowId","Replicates","Prevalence")]
wide_m_P <- spread(wide_m_P,key=Replicates,Prevalence)
wide_m_P <- separate(wide_m_P,col="MeadowId",into=c("Region","SiteCode"),sep="_",remove = FALSE)

wide_m_S <- full_m[,c("MeadowId","Replicates","Severity")]
wide_m_S <- spread(wide_m_S,key=Replicates,Severity)
wide_m_S <- separate(wide_m_S,col="MeadowId",into=c("Region","SiteCode"),sep="_",remove = FALSE)

wide_m_BA <- full_m[,c("MeadowId","Replicates","BladeArea")]
wide_m_BA <- spread(wide_m_BA,key=Replicates,BladeArea)
wide_m_BA <- separate(wide_m_BA,col="MeadowId",into=c("Region","SiteCode"),sep="_",remove = FALSE)

mP15 <- ggplot(wide_m_P,aes(x=Twenty))+
  geom_point(aes(y=Fifteen,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  theme_bw()+
  labs(title="15 blades")+
  xlab("")+
  ylab("Prevalence")+
  theme(legend.position = "")

mP10 <- ggplot(wide_m_P,aes(x=Twenty))+
  geom_point(aes(y=Ten,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  theme_bw()+
  labs(title="10 blades")+
  xlab("")+
  ylab("Prevalence")

mP5 <- ggplot(wide_m_P,aes(x=Twenty))+
  geom_point(aes(y=Five,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  #  stat_smooth(method="lm",geom="smooth",aes(y=Five))+
  theme_bw()+
  labs(title="5 blades")+
  xlab("Prevalence of n=20 blades")+
  ylab("Prevalence")+
  theme(legend.position = "")

mBA15 <- ggplot(wide_m_BA,aes(x=Twenty))+
  geom_point(aes(y=Fifteen,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  #stat_smooth(method="lm",geom="smooth",aes(y=Fifteen))+
  theme_bw()+
  xlab("")+
  labs(title="15 blades")+
  ylab("Blade Area")+
  theme(legend.position = "")

mBA10 <- ggplot(wide_m_BA,aes(x=Twenty))+
  geom_point(aes(y=Ten,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  #stat_smooth(method="lm",geom="smooth",aes(y=Fifteen))+
  theme_bw()+
  labs(title="10 blades")+
  xlab("")+
  ylab("Blade Area")

mBA5 <- ggplot(wide_m_BA,aes(x=Twenty))+
  geom_point(aes(y=Five,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  #stat_smooth(method="lm",geom="smooth",aes(y=Fifteen))+
  theme_bw()+
  xlab("Blade Area of n=20 blades")+
  ylab("Blade Area")+
  labs(title="5 blades")+
  theme(legend.position = "")

mS15 <- ggplot(wide_m_S,aes(x=Twenty))+
  geom_point(aes(y=Fifteen,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  #stat_smooth(method="lm",geom="smooth",aes(y=Fifteen))+
  theme_bw()+
  xlab("")+
  ylab("Severity")+
  theme(legend.position = "")+
  labs(title="15 blades")

mS10 <- ggplot(wide_m_S,aes(x=Twenty))+
  geom_point(aes(y=Ten,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  #stat_smooth(method="lm",geom="smooth",aes(y=Fifteen))+
  theme_bw()+
  xlab("")+
  ylab("Severity")+
  labs(title="10 blades")

mS5 <- ggplot(wide_m_S,aes(x=Twenty))+
  geom_point(aes(y=Five,color=Region))+
  geom_abline(slope=1,intercept = 0,linetype="dashed")+
  #stat_smooth(method="lm",geom="smooth",aes(y=Fifteen))+
  theme_bw()+
  xlab("Severity of n=20 blades")+
  ylab("Severity")+
  labs(title="5 blades")+
  theme(legend.position = "")
```

### Meadow level plots

Prevalence

```{r mP,echo=FALSE}
mP15/mP10/mP5
```

Severity

```{r mS,echo=FALSE}
mS15/mS10/mS5
```

Blade Area

```{r mBA,echo=FALSE}
mBA15/mBA10/mBA5
```

Overall, we could probably go down to 15 blades per transect and not lose much information. There will be more spread at transect level, but the meadow level will look the same.

Blade Area looks fine even with just 5 blades, which is nice to know since our shoot morphology uses only 5 blades/shoots per transect.